### 2025-06-06 correlation.qmd
20250606 - tweaks

The covariance is unstandardized—its metric depends on the metrics of the two variables examined in the association.
However, it can be helpful to use a standardized index of the association between the variables.
A standardized index means that the index is on a common metric and does not depend on the metrics of the two variables.
This is valuable because it allows fairly comparing the strength of associations between sets of variables with different metrics, which is meaningful for estimating the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
The [Pearson correlation coefficient](#sec-correlationExamplesPearson), $r$, is an example of a standardized index of the covariance between two variables.
We can convert a covariance to a [correlation](#sec-correlationExamplesPearson) by standardizing it; i.e., by dividing by the multiplication of the standard deviation of $x$ and the standard deviation of $y$, as in @eq-pearsonCorrelation.

@@

Correlation is a standardized index of the association between two variables.
Pearson correlation is the most common type of corrrelation.
The formula for Pearson correlation is in @eq-pearsonCorrelation.
Pearson correlation assumes that the variables are [interval](#sec-interval) or [ratio](#sec-interval) level of measurement.
If the data are [ordinal](#sec-ordinal), [Spearman correlation](#sec-correlationExamplesSpearman) is more appropriate.

>>>>

The covariance is unstandardized—its metric depends on the metrics of the two variables examined in the association.
However, it can be helpful to use a standardized index of the association between the variables.
A standardized index means that the index is on a common metric and does not depend on the metrics of the two variables.
This is valuable because it allows fairly comparing the strength of associations between sets of variables with different metrics, which is meaningful for estimating the [effect size](#sec-practicalSignificance)—i.e., the strength of the association.
The [Pearson correlation coefficient](#sec-correlationExamplesPearson), $r$, is an example of a standardized index of the covariance between two variables because it has a common metric (whose possible values range from −1 to +1, regardless of the variables examined in the association).
We can convert a covariance to a [correlation](#sec-correlationExamplesPearson) by standardizing it; i.e., by dividing by the multiplication of the standard deviation of $x$ and the standard deviation of $y$, as in @eq-pearsonCorrelation.

@@

Correlation is a standardized index of the association between two variables.
Pearson correlation is the most common type of correlation.
The formula for Pearson correlation is in @eq-pearsonCorrelation.
Pearson correlation assumes that the variables are [interval](#sec-interval) or [ratio](#sec-interval) level of measurement.
If the data are [ordinal](#sec-ordinal), [Spearman correlation](#sec-correlationExamplesSpearman) is more appropriate.
